---
title: "Prediction Assignment"
author: "AG"
date: "4 November 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load the required package

```{r load packages}
library(ggplot2)
library(dplyr)
library(readr)
library(tidyr)
library(caret)
library(parallel)
library(doParallel)
```


# Download & Read files 

```{r download files}

setwd("~/Cap/03 - Support/04 - Knowledge/30 - Analytics/17 - Data Science John Hopskins/08 - Applied Machine Learning")

url_train <- 'https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv'
url_test <- 'https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv'

destfile_train <- "pml-training.csv"
destfile_test <- "pml-testing.csv"

download.file(url = url_train, destfile = destfile_train)
download.file(url = url_test, destfile = destfile_test)

df_train <- read.csv(destfile_train)
df_test <- read.csv(destfile_test)


```

# Data exploration and preparation

```{r data exploration0}
dim(df_train)
```

The training data comprises 160 variables and 19622 observations. Running a summary of the data, suggest that there is a lot of NAs and blank values in many variables as demonstrated below.

```{r data exploration_NA}

dim(df_train)

# Helper function to identify NAs values 
f_na <- function(a){
    sum(is.na(a))
}

nacount <- apply(df_train, 2, f_na)
nacount <- as.data.frame(nacount)

ggplot(nacount, aes(nacount)) + geom_histogram(binwidth = 50)
length(nacount[nacount$nacount>0, ])

```

There is 67 variables that have many NAs. Those variables will be therefore be removed from our attempt to build a predictive model.

```{r data prep_na}

df2_train <- df_train

# Remove NA values
df2_train[, nacount > 0] <- NULL

```



```{r data preparation_na}

# Helper function to identify blank values 
f_blank <- function(a) {
    sum(a == "")
}

blankcount <- apply(df2_train, 2, f_blank)
blankcount <- as.data.frame(blankcount)

ggplot(blankcount, aes(blankcount)) + geom_histogram(binwidth = 50)
length(blankcount[blankcount>0])

```

Conlusion: There is an extra 33 variables that have many blanks. Those variables will be therefore be removed from our attempt to build a predictive model.

```{r data preparation_blank}
# Remove blank values
df2_train[, blankcount > 0] <- NULL
```


# Remove variables that does not bring info

The following variables do not seem to be useful to predict the manner in which people did the exercise so they will be removed from our attempt to build a predictive model.

```{r remove variables}
df2_train$X <- NULL
df2_train$user_name <- NULL
df2_train$raw_timestamp_part_1 <- NULL
df2_train$raw_timestamp_part_2 <- NULL
df2_train$cvtd_timestamp <- NULL
df2_train$new_window <- NULL
df2_train$num_window <- NULL
```


# Build model

Given that some varibles have a very high numbers of NAs and blanks, it is legitimate to discard them to build an initial model.

```{r build_model}

#Set parallel processing
cluster <- makeCluster(detectCores() - 1) # convention to leave 1 core for OS
registerDoParallel(cluster)

# Set training parameters
fitcontrol <- trainControl(method = "cv",
                           number = 10,
                           allowParallel = TRUE)

model1 <- train(classe ~ ., method = "rf", data = df2_train, trControl = fitcontrol)

#Stop parallel processing
stopCluster(cluster)
registerDoSEQ()

#Analyse the confusion matrix
model1$finalModel$confusion[, "class.error"]

sum(diag(model1$finalModel$confusion))/sum(model1$finalModel$confusion) * 100

pred1 <- predict(model1, newdata = df_test)

pred1




```

